---
title: "Descriptive Statistics Covid Affect Mobility Project"
author: "Jordan Alexander"
date: "11/29/2022"
output: html_document
---


# Missing Data

## Mobility Data

load packages and set working directory

```{r}
setwd("C:\\Users\\alexa877\\Documents\\UMN Courses\\Research\\Covid Location project\\RAship 2022\\Datasets")
library(dplyr)
library(lubridate)
library(data.table)
library(tidyr)
library(knitr)
```

First let's restrict to the period of interest, when we were collecting affect data. This started after 02/05/2019.
```{r}
# load data
dat2 = fread('mobility.affect.covid.11.16.22.csv', data.table =  F)
min(dat2$date[is.na(dat2$pa.av) == F])
dat.tmp = dat2[dat2$date >= ymd('2019-02-05'),]

```

Report percent of missing days per participant

```{r}
missdays = dat.tmp %>% group_by(user_id) %>% 
     summarize(first = min(date, na.rm = T),
               last = max(date, na.rm = T),
               n = n())

missdays$maxdays = missdays$last - missdays$first + 1

missdays$pctmissing = 1- missdays$n/as.numeric(missdays$maxdays)
summary(missdays)
```
The average participant is missing 39.2% of days, first quartile missing 17.6%, 3rd missing 57.3%

```{r}
hist(missdays$pctmissing)
# right skewed, and kurtotic
```

Now we'll look at the length of missing intervals

```{r}
# interval length
setorder(dat.tmp, user_id, date)
dat.tmp = dat.tmp[is.na(dat.tmp$user_id) == F,]

# flag first observation for each participant
dat.tmp$firstobs = NA
dat.tmp$firstobs[dat.tmp$user_id == lag(dat.tmp$user_id)] = F
dat.tmp$firstobs[dat.tmp$user_id != lag(dat.tmp$user_id)] = T
dat.tmp$firstobs[1] = T


# calculate interval
dat.tmp$interval = dat.tmp$date - lag(dat.tmp$date)
dat.tmp$interval[dat.tmp$firstobs == T] = NA
```


85% of days are consecutive, 99% of observations come within one week of the previous observation
```{r}
summary(as.numeric(dat.tmp$interval))

length(dat.tmp$interval[dat.tmp$interval == 1])/length(dat.tmp$interval)
# 85% of intervals are only 1 day apart
length(dat.tmp$interval[dat.tmp$interval <= 2])/length(dat.tmp$interval)
# 94% 2 days apart or less
length(dat.tmp$interval[dat.tmp$interval <= 7])/length(dat.tmp$interval)
# 99% a week apart or less
```


The average participant has about 71 intervals of missing data with an average length of 5 days. Most missing days are concentrated in one long interval of about 92 days.
```{r}
# average and maximum intervals apart
tmp = dat.tmp %>% subset(firstobs == F) %>% 
     group_by(user_id) %>% 
     summarise(mean.int = mean(interval, na.rm = T),
               median.int = median(interval, na.rm = T),
               max.int = max(interval, na.rm = T),
               int2 = sum(interval >=2))
               

mean(tmp$mean.int)
# average participant has a mean interval of 5 days, skewed by outlying participants
hist(as.numeric(tmp$mean.int))


mean(tmp$int2)
hist((tmp$int2))


mean(tmp$max.int)
#average participant has a max interval of greater than 101 days, right skewed
hist(as.numeric(tmp$max.int))
summary(as.numeric(tmp$max.int))
```


## Affect data

now we give missingness descriptives for affect data. First we limit to participants who provided at least one affect survey.

```{r}
tmp2 = fread('PANASphone.long.csv') %>% mutate(date = as_date(SUBMIT_DATE),
                                               Year_Week = paste(year(date),
                                                                 week(date),
                                                                 sep = '_'))

af.miss = data.frame()

#users = unique(tmp2$user_id)
#i = users[5]

for (i in unique(tmp2$user_id)){
     temp = subset(tmp2, user_id == i)
     temp2 = subset(dat.tmp, user_id == i)
     if(nrow(temp2) == 0){temp2 = temp}
     temp = temp[is.na(temp$date) == F,]
     temp2 = temp2[is.na(temp2$date) == F,]
     first = min(temp$date,na.rm = T)
     last = max(temp2$date, na.rm = T)
     last2 = max(temp$date, na.rm = T)
     if(last2 > last){last = last2}
     n.survey = length(unique(temp$Year_Week))
     max.survey = round(as.numeric(last - first + 1)/14,
                      digits = 0)
     pct.miss = 1 - n.survey/max.survey
     if(max.survey == 0){pct.miss = NA}
     if(pct.miss < 0){pct.miss = 0}
     
     row = data.frame(first = first,
                      last = last,
                      n.survey = n.survey,
                      max.survey = max.survey,
                      pct.miss = pct.miss)
     
     af.miss = rbind(af.miss,row)
     #print(i)
     }

af.miss = subset(af.miss, n.survey > 0)
summary(af.miss$n.survey)
summary(af.miss$pct.miss)
hist(af.miss$pct.miss)
```

The average participant is contributing 25 surveys and missing 60% of their total possible affect surveys. 3rd quartile is missing 75% of them. The distribution is left skewed, most participants are missing many surveys





# Descriptive Statistics


## Mobility


Again limiting to the period of interest after February, 2019:

```{r}
get.descriptives = function(vec){
     mean = round(mean(vec,na.rm = T),digits = 2)
     sd = round(sd(vec, na.rm = T),digits = 2)
     min = round(min(vec, na.rm = T),digits = 2)
     max = round(max(vec, na.rm = T),digits = 2)
     n.obs = length(na.omit(vec))
     
     row = data.frame(mean = mean,
                      sd = sd,
                      min = min,
                      max = max,
                      n.obs = n.obs)
     return(row)}


points = get.descriptives(dat.tmp$numpts)
distance = get.descriptives(dat.tmp$distance)
pa.av = get.descriptives(dat.tmp$pa.av)
na.av = get.descriptives(dat.tmp$na.av)
weekly.cases = get.descriptives(dat.tmp$cases.weekly)
weekly.cases.per100k = get.descriptives(dat.tmp$cases.weekly.per.100k)

# build table
table1 = rbind(points,distance,pa.av,na.av,weekly.cases,weekly.cases.per100k)
rownames(table1) = c('Daily Locations Visited',
                     'Daily Distance Travelled (km)',
                     'Average Positive Affect',
                     'Average Negative Affect',
                     'Weekly Covid-19 Cases',
                     'Weekly Covid-19 Cases (per 100k)')
names(table1) = c('Mean','SD','Minimum','Maximum','n')

table1$'Missing Cases (average participant)' = c(0.39, 0.39, 0.60, 0.60, 0, 0)
```

```{r}
kable(table1)

```



## ICCs for mobility measures

```{r}
# pts ICC
dat.tmp = dat %>% 
     filter(SVID != '') %>% 
     select(SVID, date, numpts) %>% 
     arrange(SVID, date) %>% 
     select(!c(date)) %>%
     group_by(SVID) %>%
     mutate(row = row_number()) %>%
     spread(key = row, value = numpts) %>%
     ungroup()%>%select(!c(SVID))

ICC(dat.tmp)

# distance ICC

dat.tmp = dat %>% 
     filter(SVID != '') %>% 
     select(SVID, date, distance) %>% 
     arrange(SVID, date) %>% 
     select(!c(date)) %>%
     group_by(SVID) %>%
     mutate(row = row_number()) %>%
     spread(key = row, value = distance) %>%
     ungroup()%>%select(!c(SVID))

ICC(dat.tmp)






```
